---
title: "Section IV Metaprogramming"
author: "M Hannum"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rlang)
library(lobstr)
```

# [Section IV Metaprogramming Chapter 17: Big Picture](https://adv-r.hadley.nz/metaprogramming.html)

*Metaprogramming* ... "code is data" so you can do cool things with it. "hardest topic in this book"

*Non-standard evaluation* ... 

## Code is data

```{r}
expr(10 + 100)

```


vs `enexpr()`


"Captured expressions behave like lists" - 
I've been using this without really knowing what I'm doing! Trial and error using `[[` and `$`.

## Code is a tree

*Abstract syntax tree* i.e. AST

```{r}
lobstr::ast(f1(f2(a, b), f3(1, f4(2))))
#> █─f1 
#> ├─█─f2 
#> │ ├─a 
#> │ └─b 
#> └─█─f3 
#>   ├─1 
#>   └─█─f4 
#>     └─2
```

## Code can generate code

Using `rlang::call2() to create new trees from components. Function, then arguments. 

(** Similar to `do.call`)? Except it doesn't evaluate?

```{r}
?do.call

?call2
```

```{r}
call2("f", 1, 2, 3)
#> f(1, 2, 3)
call2("+", 1, call2("*", 2, 3))
#> 1 + 2 * 3

call2("+", 1, 2)
do.call("+", args = list(1, 2))

```





