---
title: "Chapter 11 Notes"
author: "M Hannum"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pryr)
```

# [Chapter 11: Functionals](https://adv-r.hadley.nz/functionals.html)

*** Actually for this chapter we used the 2nd edition, which is chapter 9 in the second edition, since it uses functions from the purrr package. 


## Predicate functionals

I had never seen these! "`some(.x, .p)` returns `TRUE` when it sees the first `TRUE`, and `every()` returns `FALSE` when it sees the first `FALSE`.

`detect_index` also seems super useful (returns location of the first match)

`keep(.x, .p)` keeps all matching elements

9.6.3 Exercises
1. Why isn't is.na() a predicate function? What base R function is closest to being a predicate version of is.na()? 

```{r}
x <- c(1, 2, NA)

is.na(x)

anyNA(x)
```

is.na() is not a predicate function becuase it returns more than 1 TRUE or FALSE. `anyNA` is close to a predicate function because it returns a single value, saying if there are any NA in the vector. 




