---
title: "Section IV Metaprogramming"
author: "M Hannum"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rlang)
library(lobstr)
```

# [Section IV Metaprogramming Chapter 17: Big Picture](https://adv-r.hadley.nz/metaprogramming.html)

*Metaprogramming* ... "code is data" so you can do cool things with it. "hardest topic in this book"

*Non-standard evaluation* ... 

## Code is data

```{r}
expr(10 + 100)

enexpr(x)
```


vs `enexpr()`


"Captured expressions behave like lists" - 
I've been using this without really knowing what I'm doing! Trial and error using `[[` and `$`.

## Code is a tree

*Abstract syntax tree* i.e. AST

```{r}
lobstr::ast(f1(f2(a, b), f3(1, f4(2))))
#> █─f1 
#> ├─█─f2 
#> │ ├─a 
#> │ └─b 
#> └─█─f3 
#>   ├─1 
#>   └─█─f4 
#>     └─2
```

## Code can generate code

Using `rlang::call2() to create new trees from components. Function, then arguments. 

(** Similar to `do.call`)? Except it doesn't evaluate?

```{r}
?do.call

?call2
```

```{r}
call2("f", 1, 2, 3)
#> f(1, 2, 3)
call2("+", 1, call2("*", 2, 3))
#> 1 + 2 * 3


do.call("+", args = list(1, 2))

call2("+", 1, call2("*", 2, 3))

call2("@~~@", x, y)


`+`(1, `*`(2, 3))
```

```{r}
xx <- expr(x + x)
yy <- expr(y + y)

xx

expr(!!xx / !!yy)
expr(xx / yy)

expr(!!enexpr(xx)/!!enexpr(yy))
#> (x + x)/(y + y)

tt <-
```


"Avoiding paste() when generateing R code" - thoughts?

## Evaluation runs code

```{r}
eval(call2("+", 1, 2))

# set environment within eval
eval(expr(x + y), env(x = 1, y = 2))

# or set the environment outside the eval and eval looks for it
x <- 10
y <- 100
eval(expr(x + y))
```

-- What is a "data mask"

## Customizing evaluation with functions
Can do things like set the caller environment i.e. bind names to functions.

## Customising evaluation with data
*** Looking for variables in a data frame instead of an environment

```{r}
df <- data.frame(x = 1:5, y = sample(5))

eval_tidy(expr(x + y), df)

with2 <- function(df, expr) {
  a <- 1000
  eval_tidy(enexpr(expr), df)
}

a <- 10
with2(df, x + a)

with3 <- function(df, expr) {
  a <- 1000
  eval_tidy(enquo(expr), df)
}

a <- 20
with3(df, x + a)
rm(list = "a")
```

## Quosures

**Quosure** bundles an expression with an environment.

Means you should use `enquo()` instead of `enexpr()` when you use a data mask. 

# Ch 18 Expressions

## Introduction

`expr()` returns an expression, which is an object that captures the structure of the code without evaluating it.

## Abstract syntax trees

```{r}
lobstr::ast(f(x, "y", 1))
#> █─f 
#> ├─x 
#> ├─"y" 
#> └─1
```

ASTs describe evaluation order, nesting of function calls

### Non-code components
 Don't capture whitespace/comments except to distinguish `<-` and `<` `-`








