---
title: "Section IV Metaprogramming"
author: "M Hannum"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rlang)
library(lobstr)
```

# [Section IV Metaprogramming Chapter 17: Big Picture](https://adv-r.hadley.nz/metaprogramming.html)

*Metaprogramming* ... "code is data" so you can do cool things with it. "hardest topic in this book"

*Non-standard evaluation* ... 

## Code is data

```{r}
expr(10 + 100)

enexpr(x)
```


vs `enexpr()`


"Captured expressions behave like lists" - 
I've been using this without really knowing what I'm doing! Trial and error using `[[` and `$`.

## Code is a tree

*Abstract syntax tree* i.e. AST

```{r}
lobstr::ast(f1(f2(a, b), f3(1, f4(2))))
#> █─f1 
#> ├─█─f2 
#> │ ├─a 
#> │ └─b 
#> └─█─f3 
#>   ├─1 
#>   └─█─f4 
#>     └─2
```

## Code can generate code

Using `rlang::call2() to create new trees from components. Function, then arguments. 

(** Similar to `do.call`)? Except it doesn't evaluate?

```{r}
?do.call

?call2
```

```{r}
call2("f", 1, 2, 3)
#> f(1, 2, 3)
call2("+", 1, call2("*", 2, 3))
#> 1 + 2 * 3


do.call("+", args = list(1, 2))

call2("+", 1, call2("*", 2, 3))

call2("@~~@", x, y)


`+`(1, `*`(2, 3))
```

```{r}
xx <- expr(x + x)
yy <- expr(y + y)

xx

expr(!!xx / !!yy)
expr(xx / yy)

expr(!!enexpr(xx)/!!enexpr(yy))
#> (x + x)/(y + y)

tt <-
```


"Avoiding paste() when generateing R code" - thoughts?

## Evaluation runs code

```{r}
eval(call2("+", 1, 2))

# set environment within eval
eval(expr(x + y), env(x = 1, y = 2))

# or set the environment outside the eval and eval looks for it
x <- 10
y <- 100
eval(expr(x + y))
```

-- What is a "data mask"

## Customizing evaluation with functions
Can do things like set the caller environment i.e. bind names to functions.

## Customising evaluation with data
*** Looking for variables in a data frame instead of an environment

```{r}
df <- data.frame(x = 1:5, y = sample(5))

eval_tidy(expr(x + y), df)

with2 <- function(df, expr) {
  a <- 1000
  eval_tidy(enexpr(expr), df)
}

a <- 10
with2(df, x + a)

with3 <- function(df, expr) {
  a <- 1000
  eval_tidy(enquo(expr), df)
}

a <- 20
with3(df, x + a)
rm(list = "a")
```

## Quosures

**Quosure** bundles an expression with an environment.

Means you should use `enquo()` instead of `enexpr()` when you use a data mask. 

# Ch 18 Expressions


## Introduction

`expr()` returns an expression, which is an object that captures the structure of the code without evaluating it.

## Abstract syntax trees

```{r}
lobstr::ast(f(x, "y", 1))
#> █─f 
#> ├─x 
#> ├─"y" 
#> └─1
```

ASTs describe evaluation order, nesting of function calls

### Non-code components
ASTs don't capture whitespace/comments except to distinguish `<-` and `<` `-`

### Infix calls

"Infix" notation is where operators are "infixed" between operands (e.g. y <- a + b). 

### Exercises

1. Reconstruct code represented by trees
```{r}
lobstr::ast(f(g(h())))

lobstr::ast(1+2+3)

lobstr::ast((x+y)*z)
```

2. Draw trees by hand and check answers with `lobstr::ast()`

```{r}
lobstr::ast(f(g(h(i(1, 2, 3)))))
lobstr::ast(f(1, g(2, h(3, i()))))
lobstr::ast(f(g(1, 2), h(3, i(4, 5))))
```

3. What's happening with the ASTs below?

    1. variable x added to variable y

```{r}
lobstr::ast(`x` + `y`)
#> █─`+` 
#> ├─x 
#> └─y
lobstr::ast(x + y)
```

    2. Variable x raised to power of y. Per documentation, "** is translated in the parser to ^, but this was undocumented for many years."
    
```{r}

lobstr::ast(x ** y)
#> █─`^` 
#> ├─x 
#> └─y
```

    3. Parser prefers '<-' over '->'
```{r}
lobstr::ast(1 -> x)
#> █─`<-` 
#> ├─x 
#> └─1
```

4. What is special about the AST below?

`srcref` is short for source reference. It's not specified in the expression within the ast 
```{r}
lobstr::ast(function(x = 1, y = 2) {})
#> █─`function` 
#> ├─█─x = 1 
#> │ └─y = 2 
#> ├─█─`{` 
#> └─<inline srcref>


```


## Expressions

*Expression* = "any member of the set of base types created by parsing code: constant scalars, symbols, call objects, and pairlists. 

### Constants

Constant is either NULL or length-1 atomic vector
```{r}
identical(expr(TRUE), TRUE)
#> [1] TRUE
identical(expr(1), 1)
#> [1] TRUE
identical(expr(2L), 2L)
#> [1] TRUE
identical(expr("x"), "x")
#> [1] TRUE
```

### Symbols

Symbol represents the name of an object

***`rlang::sym()` seems it could be useful?

```{r}
sym("x")

syms(c("x", "y"))
```

Can go back with `as.character()` or `rlang::as_string()`. 

### Calls

Call object is a special type of list - where first element is function call. 

```{r}
lobstr::ast(read.table("important.csv", row.names = FALSE))
#> █─read.table 
#> ├─"important.csv" 
#> └─row.names = FALSE
x <- expr(read.table("important.csv", row.names = FALSE))
```


#### Subsetting
Can use standard subsetting on calls because they behave like lists! 


* can we talk about the differences between the below? 

```{r}
x[[1]]

x[[2]]

x[[3]]

is.expression(x[-1])
is.expression(x)
class(x)

x[1]

x[2]

as.list(x)

lobstr::ast(!!x)

x$sep <- ""
```

*** I love that calls can be modified like lists



#### Function position

Possible that first element of call object (i.e. "function position") isn't a symbol, might have to be retrieved. 

#### Constructing

Simple expressions can be constructed using `rlang::call2()`. 

```{r}
# first argument name of 
call2("mean", x = expr(x), na.rm = TRUE)
#> mean(x = x, na.rm = TRUE)
call2(expr(base::mean), x = expr(x), na.rm = TRUE)
#> base::mean(x = x, na.rm = TRUE)
```


### Exercises

1. Which two of the six types of atomic vector can't appear in an expression? 

A: Have to be atomic of length 1. e.g. can't be complex atomics or raws. 
Only 4 allowed are: int, num, chr, lgl.

2. What happens when you subset a call object to remove the first element? 

```{r}
y <- expr(read.csv("foo.csv", header = TRUE))[-1]

is_call(y)
is_expression(y)

y$sep <- ""
```

Object is no longer an expression with out the function call. The first element is now the 2nd element of the original. 

3. Describe the differences: 

```{r}
x <- 1:10

call2(median, x, na.rm = TRUE)
eval(call2(median, x, na.rm = TRUE))

call2(expr(median), x, na.rm = TRUE)
eval(call2(expr(median), x, na.rm = TRUE))

call2(median, expr(x), na.rm = TRUE)
eval(call2(median, expr(x), na.rm = TRUE))


call2(expr(median), expr(x), na.rm = TRUE)
eval(call2(expr(median), expr(x), na.rm = TRUE))
```

All are evaluated to the same response, but differ on when `x` and `median()` are evaluated. 

4. `rlang::call_standardise()` doesn't work so well for `mean()` since `mean()` uses the `...` argument which is difficult to standardize. 


```{r}
library(rlang)

call_standardise(quote(mean(1:10, na.rm = TRUE)))
#> mean(x = 1:10, na.rm = TRUE)
call_standardise(quote(mean(n = T, 1:10)))
#> mean(x = 1:10, n = T)
call_standardise(quote(mean(x = 1:10, , TRUE)))
#> mean(x = 1:10, , TRUE)

call_standardise(quote(mean.default(x = 1:10, , TRUE)))


```

Try to use `mean.default()` instead. 

5. Why doesn't this code make sense? 

```{r}
x <- expr(foo(x = 1))
names(x) <- c("x", "")


x

names(x) <- c("", "x")

x
```

Doesn't make sense to name a function call?

6. Construct the expression `if(x > 1) "a" else "b"` using multiple `call2()`. 

```{r}
x = 10
# first attempt
call2("ifelse", expr(x > 1), "a", "b")


# second attempt
#call2("if", call2(">", x, 1), "a", "b")

# finally
call2("if", call2(">", sym("x"), 1), "a", "b")

# check with eval
x = 10
eval(call2("if", call2(">", sym("x"), 1), "a", "b"))

x = 0
eval(call2("if", call2(">", sym("x"), 1), "a", "b"))
```

## Parsing and grammar

"Parsing" = "process by which a computer language takes a string and constructs an expression" by rules of that languages "grammar".

### Operator precedence

Just like in school maths :) 

EXCEPT the `!` has low precedence!

```{r}
# use this to see operator description and precedence
?Syntax
```

* Note parenthesis as a call in AST

```{r}
lobstr::ast((1 + 2) * 3)
#> █─`*` 
#> ├─█─`(` 
#> │ └─█─`+` 
#> │   ├─1 
#> │   └─2 
#> └─3
```

### Associativity

Most operators are left associative except for exponentiation and assignment. 

### Parsing and deparsing
What to do when you have code stored in a string... and want to parse it yourself! 

```{r}
x1 <- "y <- x + 10"
x1
#> [1] "y <- x + 10"
is.call(x1)
#> [1] FALSE

x2 <- rlang::parse_expr(x1)
x2
#> y <- x + 10
is.call(x2)
#> [1] TRUE
```

* Use `rlang::parse_exprs()` for a list of expressions. 

### Exercises
1. R uses parentheses in two slightly different ways: compare and contrast

```{r}
f((1))
`(`(1 + 1)
```

Parenthesis can be part of a function or just part of syntax.. 

When used in `f()` the parenthesis is part of syntax, while the inner one around 1 (`(1)`) is a function. 

I don't know exactly why the second version only calls one `(` function, see my second example. 
```{r}
ast(f((1)))


ast(`(`(1 + 1))
#equivalent to this for some reason??
ast((1 + 1))

# and here we have 2 `(` functions in the AST tree
ast(((1 + 1)))
```

2. `=` can be used 2 ways. show an example. 

```{r}
# assignment and as part of syntax of function calls

df = data.frame(x = 2)

```

3. Does `-2^2` yield 4 or -4?

Yields -4 because of syntax precedence. `^` takes precedence over `-`
```{r}
-2^2

?Syntax
```

4. Why does `!1 + !1` return what it does? 

```{r}
!1 + !1

ast(!1 + !1)

!1

1 + !1
```

Unexpected (maybe) evaluation

5. Why does `x1 <- x2 <- x3 <- 0 ` work? Describe two reasons

```{r}
x1 <- x2 <- x3 <- 0 
```

Assignment goes right to left (as does `^`). 

Had to look up other reason: that `<-` "invisibly returns the value on the right hand side". 

```{r}
(x3 <- 0)
```

6. Compare ASTs of x + y %+% z and x ^ y %+% z. 

```{r}
ast(x + y %+% z)

!%in%
  
  !(  %in% )

?`%+%`

ast(x ^ y %+% z)
```

Custom infix `%+%` seems to have precedence over `+` but not over `^`. 
















