---
title: "Chapter 3 Notes"
author: "M Hannum"
date: "Sys.Date()"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 3: Subsetting 

Three subsetting operators:

Six types of subsetting:


## Examples for quiz

1. Result of subsetting a vector with positive integers, negative integers, logical vector, or a character vector
```{r}

df <- data.frame(x = c(1:3), y = c("a", "b", "c"))

#keep elements
df[1,]

#drop elements
df[-1,]

#keep TRUE elements
df[c(TRUE, TRUE, FALSE),]

#keep elements with matching names
df[,"y"]

```

2. Difference between [, [[, and $ when applied to a list

```{r}
df2 <- data.frame(z = c(4:6))
l1 <- list(df = df, df2 = df2)

#returns a list
str(l1["df"])

#returns element in a list
str(l1[["df"]])

#equivalent to [[]]
str(l1$df)

```

3. Use `drop = FALSE` when subsetting inside a function - WHY?

4. What does `x[] <- 0` do to a matrix `x`?

```{r}
x <- c(1:12)
dim(x) <- c(3,4)

x

#Make subset element of matrix 0 
x[1,1] <- 0

x

#makes every element of the matrix 0
x[] <- 0 

x
```

5. How to use a named vector to relabel categorical variables


```{r}
x1 <- c(x = 1, y = 2, z = 3)

c(x = 1, y = 2, z = 3)[c("x", "y", "z")]

df3 <- cbind(df, df2)
names(df3)

#use named character vector as lookup
names(df3)[x1[c("x", "y")]] <- x1[c("x", "y")]

names(df3)

```

