---
title: "Chapter 6 Notes"
author: "M Hannum"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pryr)
```

# [Chapter 6: Functions](http://adv-r.had.co.nz/Functions.html)

**Opening quiz**  

1. 3 components of a function: body, arguments, environment  
2. 11 (comment: generally should functions within functions take input from the surrounding function? Haven't used the double ()() before)  


```{r}
x <- 10
f1 <- function(x) {
  function() {
    x + 10
  }
}
f1(1)()

```

3. `1 + (2 * 3)`  

4. 

```{r}

mean(c(1:10, NA), na.rm = TRUE)
```

5. `b` argument is never used in function so it does not stop or produce error message. 

```{r}
f2 <- function(a, b) {
  a * 10
}
f2(10, stop("This is an error!"))
```

6. infix function is binary operator function

## Function components

**6.1.2 Exercises**  
1. Can use `class()` to tell if an object is a function. Can also use `is.function()` and `is.primitive()` for a logical test on if object is a function or primitive function.  

2.  
a. The `scan()` function has the most arguments (22)  

```{r}
# make list of all functions in base package
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)

# find which base function has most arguments

# create a named vector of lengths
arg_length <- purrr::map_int(funs, function(x)(length(formals(x))))

#Get name of function with most arguments
max(arg_length)
names(which(arg_length == max(arg_length)))

```

b. 
```{r}
no_args <- names(which(arg_length == 0))
head(no_args)
length(no_args)


prims <- Filter(is.primitive, objs)
length(prims)

arg_length_prims <- purrr::map_int(prims, function(x)(length(formals(x))))

non_prim_no_args <- no_args[!(no_args %in% names(prims))]

non_prim_no_args
```



3. 

