---
title: "Chapter 6 Notes"
author: "Daniel Sjoberg"
date: "January 21, 2019"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Quiz

#### 1. What are the three components of a function?

Hmmm, call, body, inputs.

*Correct answer: body, arguments, and environment.*

#### 2. What does the following code return?

```{r remedy001, eval = F}
x <- 10
f1 <- function(x) {
  function() {
    x + 10
  }
}
f1(1)()
```

`f1(1)() = 11`

#### 3. How would you more typically write this code?

```{r remedy002, eval = F}
`+`(1, `*`(2, 3))
```

`1 + 2 * 3`

#### 4. How could you make this call easier to read?

```{r remedy003, eval = F}
mean(, TRUE, x = c(1:10, NA))
```

```{r remedy004, eval = F}
mean(c(1:10, NA), na.rm = TRUE)
```

#### 5. Does the following function throw an error when called? Why/why not?

```{r remedy005, eval = F}
f2 <- function(a, b) {
  a * 10
}
f2(10, stop("This is an error!"))
```

No, because `b` is never evaluated.

#### 6. What is an infix function? How do you write it? What’s a replacement function? How do you write it?

The R functions where the inputs are on each side of the operator, e.g. LHS %...% RHS.

#### 7. What function do you use to ensure that a cleanup action occurs regardless of how a function terminates?

IDK  
*Correct Answer: You use `on.exit()`* 

# Function components
## Exercises

#### 1. What function allows you to tell if an object is a function? What function allows you to tell if a function is a primitive function?

`is.function()` and `is.primitive()`

The `formals()`, `body()`, and `environment()` functions can assist with both of these.  These will all be `NULL` for primitive functions.

#### 2. This code makes a list of all functions in the base package.

```{r remedy006, eval = T}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
```

Use it to answer the following questions:
a) Which base function has the most arguments?

```{r, eval = T}
arg_n <- map_dbl(funs, ~ formals(.x) %>% length()) 
arg_n %>%
  sort(decreasing = TRUE) %>%
  head()
```

b) How many base functions have no arguments? What’s special about those functions?

```{r, eval = T}
arg_0 <- arg_n[arg_n == 0]
length(arg_0)
head(arg_0)
```

Many functions with 0 arguments are operators.

c) How could you adapt the code to find all primitive functions? 

```{r, eval = T}
objs <- mget(ls("package:base"), inherits = TRUE)
funs_primitive <- Filter(is.primitive, objs)
length(funs_primitive)
```

#### 3. What are the three important components of a function?

body, formals, and environment

#### 4. When does printing a function not show what environment it was created in?

When it is a primitive funciton

# Lexical scoping

## Exercises
#### 1. What does the following code return? Why? What does each of the three c’s mean?

```{r, eval = T}
c <- 10
c(c = c)
```

Named vector with value 10 and name c.  R scoping allows for differnet types of objects/functions to co-exist with the same name.

#### 2. What are the four principles that govern how R looks for values?

IDK

#### 3. What does the following function return? Make a prediction before running the code yourself.

```{r, eval = T}
f <- function(x) {
  f <- function(x) {
    f <- function(x) {
      x ^ 2
    }
    f(x) + 1
  }
  f(x) * 2
}
# 202?
f(10)

```

# Function arguments
## Exercises
#### 1. Clarify the following list of odd function calls:

```{r, eval = T}
set.seed(456)
x <- sample(replace = TRUE, 20, x = c(1:10, NA))
y <- runif(min = 0, max = 1, 20)
cor(m = "k", y = y, u = "p", x = x)

set.seed(456)
x <- sample(c(1:10, NA), size = 20, replace = TRUE)
y <- runif(20)
cor(x, y, use = "pairwise.complete.obs", method = "kendall")
```

#### 2. What does this function return? Why? Which principle does it illustrate?

IDK!

```{r, eval = T}
f1 <- function(x = {y <- 1; 2}, y = 0) {
  x + y
}
f1()
```

#### 3. What does this function return? Why? Which principle does it illustrate?

```{r, eval = T}
f2 <- function(x = z) {
  z <- 100
  x
}
f2()
```

100, it's lazy evaluation.








