---
title: "Chapter 6 Notes"
author: "Daniel Sjoberg"
date: "January 21, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Quiz

#### 1. What are the three components of a function?

Hmmm, call, body, inputs.

*Correct answer: body, arguments, and environment.*

#### 2. What does the following code return?

```{r remedy001, eval = F}
x <- 10
f1 <- function(x) {
  function() {
    x + 10
  }
}
f1(1)()
```

`f1(1)() = 11`

#### 3. How would you more typically write this code?

```{r remedy002, eval = F}
`+`(1, `*`(2, 3))
```

`1 + 2 * 3`

#### 4. How could you make this call easier to read?

```{r remedy003, eval = F}
mean(, TRUE, x = c(1:10, NA))
```

```{r remedy004, eval = F}
mean(c(1:10, NA), na.rm = TRUE)
```

#### 5. Does the following function throw an error when called? Why/why not?

```{r remedy005, eval = F}
f2 <- function(a, b) {
  a * 10
}
f2(10, stop("This is an error!"))
```

No, because `b` is never evaluated.

#### 6. What is an infix function? How do you write it? What’s a replacement function? How do you write it?

The R functions where the inputs are on each side of the operator, e.g. LHS %...% RHS.

#### 7. What function do you use to ensure that a cleanup action occurs regardless of how a function terminates?

IDK  
*Correct Answer: You use `on.exit()`* 