---
title: "Chapter 8 Notes"
author: "M Hannum"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pryr)
```

# [Chapter 9: Debugging](http://adv-r.had.co.nz/Exceptions-Debugging.html)

Defensive programming principles: 

* Check inputs are correct  
* Avoid non-standard evaluation  
* Avoid functions that can return different types of output

**Opening quiz**  

1. How can you find out where an error occurred? 

Look at console results or use traceback.

2. How does `browser()` do? 

??

3. What function do you use to ignore errors in a block of code? 

`tryCatch()`. Can also nest it within functions. 

4. Why might you want to create an error with a custom S3 class? 

??

```{r}
f <- function(a) g(a)
g <- function(b) h(b)
h <- function(c) i(c)
i <- function(d) "a" + d
f(10)
```

Exercises:

What is advantage of withCallingHandlers()?
```{r}
condition <- function(subclass, message, call = sys.call(-1), ...) {
  structure(
    class = c(subclass, "condition"),
    list(message = message, call = call),
    ...
  )
}
is.condition <- function(x) inherits(x, "condition")
e <- condition(c("my_error", "error"), "This is an error")
message2error <- function(code) {
  withCallingHandlers(code, message = function(e) stop(e))
}
message2error2 <- function(code) {
  tryCatch(code, message = function(e) stop(e))
}

message2error(stop(e))
traceback()
message2error2(stop(e))
traceback()
```
Fewer lines in 1st traceback? 

# Defensive programming

1. Strict input: look into `assertthat::stopifnot()`

2. Avoid functions that use non-standard evaluation (e.g. subset, transform, with)

3. Avoid functions that return different types of output depening on input. Specifically: `sapply()` and `[`

## Exercises
The goal of the col_means() function defined below is to compute the means of all numeric columns in a data frame.
```{r}
col_means <- function(df) {
  numeric <- sapply(df, is.numeric)
  numeric_cols <- df[, numeric]

  data.frame(lapply(numeric_cols, mean))
}
```


However, the function is not robust to unusual inputs. Look at the following results, decide which ones are incorrect, and modify col_means() to be more robust. (Hint: there are two function calls in col_means() that are particularly prone to problems.)

```{r}
col_means(mtcars)
col_means(mtcars[, 0])
col_means(mtcars[0, ])
col_means(mtcars[, "mpg", drop = F])
col_means(1:10)
col_means(as.matrix(mtcars))
col_means(as.list(mtcars))
mtcars
mtcars2 <- mtcars
mtcars2[-1] <- lapply(mtcars2[-1], as.character)
col_means(mtcars2)
```



